---
title: "Initial Korean covid 19"
author: "Matt Diaz"
date: "Spring 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?


# Introduction 

$H_0$: There is no difference in death rate based on gender. 

$H_A$: There is a difference in death rate based on gender. 



# Methods
Variables for this study are Sex and result. 

We want to find out if result depends on sex. Result is a categorical variable, sex is a categorical variable. 

So we use the Cat ~ Cat helper file. 

# Results 


## Graphical results - use ggplot commands to do this  


```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
  
This chart shows me that of all those who have dies, 60% of those were males and about 40% of those who dies are females. 


## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

This table shows the average male deaths are higher than female deaths at 64 deaths vs female death 36.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi-squared test p-value is 0.0012 which is less than 0.05 so we will reject the null hypothesis and say that there is a difference between death rate based on sex. In conclusion, these results indicate a statistically significant association between the variables with an estimated odds ratio of 2.3 and a 95% confidence interval ranging from 1.4 to 3.9. 

## Conclusions

The original question was was "is there a difference in death rate in the Korean data set based on Sex?

The plot indicates that male death is higher on average being 60% and female average below 40%

The numerical result shows the average male deaths are higher than female deaths at 64 deaths vs female death 36. We were able to conclude that there is a significant death rate difference due to Sex.

